---
title: "Response Summary for Problem Set 4"
author: "William W. Chiang"
date: "January 5, 2016"
output: html_document
---

Diamonds Data

Notes: Prices of 50,000 round cut diamonds

Description

A dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:

Usage

data(diamonds) Format

A data frame with 53940 rows and 10 variables

Details

price. price in US dollars (\$326–\$18,823)

carat. weight of the diamond (0.2–5.01)

cut. quality of the cut (Fair, Good, Very Good, Premium, Ideal)

colour. diamond colour, from J (worst) to D (best)

clarity. a measurement of how clear the diamond is (I1 (worst), SI1, SI2, VS1, VS2, VVS1, VVS2, IF (best))

x. length in mm (0–10.74)

y. width in mm (0–58.9)

z. depth in mm (0–31.8)

depth. total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79)

table. width of top of diamond relative to widest point (43–95)

***
### Load ggplot2 package and diamonds dataset
```{r}
library(ggplot2)
data(diamonds)
```
***
### Your first task is to create a scatterplot of price vs x. using the ggplot syntax. You can also embed plots, for example:

```{r}
ggplot(aes(x = price, y = x), data = diamonds) +
  geom_point()
```

There is a positive correlation between price and x (diamond length in mm). Furthermore, the rate of x increase is not constant between the low and high end of the price range. Diamond length increases at a greater rate for incremental price increments between 326 (price minimum, summary(diamonds)) and 2500 than prices 2501 and above.

***
### Load the alr3 package for calculating correlation coefficients
```{r}
library(alr3)
```

***
### What is the correlation between price and x?
```{r}
cor.test(x = diamonds$price, y = diamonds$x)
```

### What is the correlation between price and y?
```{r}
cor.test(x = diamonds$price, y = diamonds$y)
```

### What is the correlation between price and z?
```{r}
cor.test(x = diamonds$price, y = diamonds$z)
```

***
### Create a simple scatter plot of price vs depth
```{r}
ggplot(aes(x = price, y = depth), data = diamonds) +
  geom_point()
```

***
### Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units
```{r}
ggplot(aes(x = depth, y = price), data = diamonds) +
     geom_point(alpha = 1/100) +
     scale_x_continuous(breaks = seq(0, 100, 2)) +
     labs(x = "Depth (%)", y = "Price (USD)", title = "Diamond Price vs. Depth") +
     coord_cartesian(xlim = c(55,65))
```

Based on the scatterplot above, most diamonds are between 59 to 63 percent in depth.

***
### The correlation betwen depth and price can be determine from the r code below.
```{r}
cor.test(x = diamonds$depth, y = diamonds$price)
```

Result: No correlation between price and depth. Depth should not be used to predict the price of diamond because r was shown to be -0.01, which in absolute terms, is very much below the r = 0.8 significant point.

***
### Create a scatterplot of price vs carat and omit the top 1% of price and carat values
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point() +
  scale_x_continuous(limits = c(0, quantile(diamonds$carat, 0.99))) +
  scale_y_continuous(limits = c(0, quantile(diamonds$price, 0.99))) +
  labs(title = "Diamond Price vs. Mass", x = "Mass (carats)", y = "Price (USD)")  
```

***
### Create a new variable for volume in the diamonds data frame
```{r}
diamonds$volume <- with(diamonds, x * y * z)
```

### Create a scatterplot of price vs. volume (x * y * z)
```{r}
ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point()
```

Judging the price versus volume scatterplot, there is significant correlation between these two features.
***
### Find the correlation between price and volume, excluding diamonds with volumes of 0 or greater than or equal to 800
```{r}
with(diamonds[diamonds$volume > 0 & diamonds$volume < 800, ], cor(volume, price))
```

***
### Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot.
```{r}
diamonds_subset <- subset(diamonds, volume >0 & volume <= 800)

ggplot(aes(x = volume, y = price), data = diamonds_subset) +
    geom_point(alpha = 1/100) +
    geom_smooth(method = "lm", colour = "red") +
    labs(title = "Price vs. Volume", x = "Volume", y = "Price (USD)")
```

***
### Use the function dplyr package to create a new data frame containing info on diamonds by clarity, and name the data fram diamondsByClarity
```{r}
library(dplyr)

diamondsByClarity <- summarise(group_by(diamonds, clarity),
                               mean_price = mean(price),
                               median_price = median(price),
                               min_price = min(price),
                               max_price = max(price),
                               n = n()
                              )

diamondsByClarity
```

***
### Bar charts of mean price
```{r}
library(gridExtra)

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x = clarity, y = mean_price), data = diamonds_mp_by_clarity) +
  geom_bar(stat = "identity")

p2 <- ggplot(aes(x = color, y = mean_price), data = diamonds_mp_by_color) +
  geom_bar(stat = "identity")

grid.arrange(p1,p2, ncol = 1)
```

Results: Mean price tend to be higher for poorer clarity and color of diamonds!

***
